---
title: "Police Sentiment Scores | City of Chicago Dataset Project"
author: "Briggy12"
date: "2025.8.7"
categories: [data_science,general]
---

## My Project

Quick Links:

-   data.gov page: <https://catalog.data.gov/dataset/police-sentiment-scores>
    -   This is where I originally retrieved the data set from.
-   Original posting from the Chicago Data Portal: <https://data.cityofchicago.org/Public-Safety/Police-Sentiment-Scores/28me-84fj/about_data>
    -   Here's some more information on the data and explanation of variables.
-   Chicago Police Department's Sentiment Dashboard: <https://www.chicagopolice.org/statistics-data/data-dashboards/sentiment-dashboard/>
    -   Here's CPD's report on the data. On the site, they've also posted some breakdowns and graphic models of their own on the same data.

### Introduction

Over the summer, I attended a Title III/MSIEP grant workshop that focused on data science and analytics using R. As part of the workshop, we were asked to find a data set and ask a overarching question that could be solved through modeling the data, using the skills we learned in R.

### Explanation of Data/Methods

When looking for a data set, I wanted to dive into something more political and within my interests. I found [this data set](https://catalog.data.gov/dataset/police-sentiment-scores) on data.gov (<https://catalog.data.gov/dataset/police-sentiment-scores>) which showed data on the participants' trust of the police, their overall personal measurement of their safety (i.e. "how safe do you feel?"), and how much do you feel the police listen to your area and take that feedback into account.

Participants were asked to rate these factors on a scale from 1-10. Each question also was tied to the demographics of participants, concerning race, age, sex, education level (as in high school diploma or less, some college, and high ed degree), and income.

The survey was broadcasted through advertisements and was self-reporting and optional.

### Research Question

My main question was "which race trusts police/CPD the most?" so much of my models focus on that and related breakdowns of the data, but I wrote code to explore some other statistics as well.

### EDA (Exploratory Data Analysis)

Below, you'll find the code that I ran in R to create models from this data set. I used the package `ggplot2` to create models of the data set.

This first block here is breaking up each rating by race. The races we have to pick from on the data set are African-American (Black), Asian-American, Hispanic, and White. In these graphs, I ran the same code to see how much each race trusts the police as reported in the survey.

```{r}
#| fig-height: 3
#| fig-width: 5
install.packages("ggplot2")
library(ggplot2)

PSS <- read.csv("Police_Sentiment_Scores.csv")

ggplot(
  data=PSS, aes(x=T_RACE_AFRICAN_AMERICAN)
  ) + geom_density(
    color="blue", fill="blue", alpha=.5
    )

ggplot(
  data=PSS, aes(x=T_RACE_ASIAN_AMERICAN)
  ) + geom_density(
    color="green", fill="green", alpha=.5
    )

ggplot(
  data=PSS, aes(x=T_RACE_HISPANIC)
  ) + geom_density(
    color="yellow", fill="yellow", alpha=.5
    )

ggplot(
  data=PSS, aes(x=T_RACE_WHITE)
  ) + geom_density(
    color="red", fill="red", alpha=.5
    )
```

On their own though, these tell a lot about the trust level broken down by race. I'm interest in the comparison, though. Below are two graphs; one of all of the trust values by race combined and the other is a comparison between the Black trust values versus the White trust values.

```{r}
#| fig-height: 8
install.packages("ggplot2")
library(ggplot2)

PSS <- read.csv("Police_Sentiment_Scores.csv")

# graph of all races' trust values, stacked
ggplot(
  data = PSS,
  mapping = aes(y = AREA)
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_AFRICAN_AMERICAN),
  color = "blue", fill="blue", alpha = 0.3
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_ASIAN_AMERICAN),
  color = "green", fill="green", alpha = 0.3
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_HISPANIC),
  color = "yellow", fill="yellow", alpha = 0.3
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_WHITE),
  color = "red", fill="red", alpha = 0.3
)+ labs(
    title = "Trust Level for Police",
    subtitle = "Trust level based on race demographic",
    x = "Trust Level (1-100)",
    y = "Area",
    )
```

```{r}
#| fig-height: 6
install.packages("ggplot2")
library(ggplot2)

PSS <- read.csv("Police_Sentiment_Scores.csv")

# graph of Black versus White trust values
ggplot(
  data = PSS,
  mapping = aes(y = AREA)
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_AFRICAN_AMERICAN),
  color = "blue", fill="blue", alpha = 0.3
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_WHITE),
  color = "red", fill="red", alpha = 0.3
)+ labs(
    title = "Trust Level for Police",
    subtitle = "Trust level based on race demographic, only African American and White",
    x = "Trust Level (1-100)",
    y = "Area",
    )
```

### Conclusion

The results for this EDA were about what I expected. Given the condition of our systems of law enforcement, as well as the contemporary political culture, it makes sense to see how different races view the police and the general trends they follow.

What I was surprised to see is how spread out or concentrated certain results were, particularly when broken down by area. For example, "area 4"; meaning Austin, Garfield Park, North and South Lawndale; are much more spread out in their trust of the police, showing more opinion disparity. Whereas the North Side as a whole is a lot more uniform in the populations' overall view of the police.

Something I think is hard to notice (and in future re-tries of this project can be addressed) but also interesting is that some races had certain boundaries for their scoring of trust. The minimum value that Black participants as a whole gave was lower than that of the White participants, and vice versa for the upper end.
