---
title: "Police Sentiment Scores | City of Chicago Dataset Project"
author: "Briggy12"
date: "2025.8.7"
categories: [data_science,general]
---

## My Project

Quick Links:

-   [data.gov](ttps://catalog.data.gov/dataset/police-sentiment-scores "Links to data.gov's webpage featuring the dataset used for this project") page (<https://catalog.data.gov/dataset/police-sentiment-scores>)
    -   This is where I originally retrieved the data set from.
-   Original posting from the [Chicago Data Portal](https://data.cityofchicago.org/Public-Safety/Police-Sentiment-Scores/28me-84fj/about_data) (<https://data.cityofchicago.org/Public-Safety/Police-Sentiment-Scores/28me-84fj/about_data>)
    -   Here's some more information on the data and explanation of variables.
-   Chicago Police Department's [Sentiment Dashboard](https://www.chicagopolice.org/statistics-data/data-dashboards/sentiment-dashboard/) (<https://www.chicagopolice.org/statistics-data/data-dashboards/sentiment-dashboard/>)
    -   Here's CPD's report on the data. On the site, they've also posted some breakdowns and graphic models of their own on the same data.

### Introduction

and then here I'll give a description of the project

### Explanation of Data/Methods

When looking for a dataset, I wanted to dive into something more political and within my interests. I found this dataset on [data.gov](ttps://catalog.data.gov/dataset/police-sentiment-scores "Links to data.gov's webpage featuring the dataset used for this project") ([https://catalog.data.gov/dataset/police-sentiment-scores](ttps://catalog.data.gov/dataset/police-sentiment-scores)) which showed data on the "trust" of participants with the police. The "trust" was a rating of 1-10 from a variety of demographics spanning 22 districts within Chicago.

Mention interest and also survey methodology

### Research Questions

-   
-   What area is most trustworthy of CPD?
-   f

### EDA (Exploratory Data Analysis)

Below, you'll find the code that I ran in R to create models from this data set.

```{r}
install.packages("ggplot2")
library(ggplot2)

PSS <- read.csv("Police_Sentiment_Scores.csv")
print()

#lets make a general count of what participants we have?


# vv for these down here: change "N/A" to something in AREA and choose a different  color palette
ggplot(data = PSS, aes(x=T_SEX_MALE, color=AREA, fill=AREA)) + geom_bar(position="fill")
ggplot(data = PSS, aes(x=T_SEX_FEMALE, color=AREA, fill=AREA)) + geom_bar(position="fill")

# ehhh this one kinda sucks because they're independant variables
ggplot(data = PSS, aes(x=S_RACE_WHITE, y=AREA, color=AREA, fill=AREA)) + geom_point(size=.5, alpha=.5, position="jitter")


## maybe if I want to make a "map" of where people live, try sorting a count-based graph with each of the race categorizations? Therefore omiting their values and just counting the occurances. Could probably make a new variable using the count() function

#use tidyverse gather()

```

So let's check by area.

In the dataset, there's 9 categorizations of the reported area (each of these are better explained in [CPD's Sentiment Dashboard](https://www.chicagopolice.org/statistics-data/data-dashboards/sentiment-dashboard/)):

1.  **"N/A"**, or no area reported
2.  **"Area 1"**, meaning McKinley Park, Chicago Lawn, Ashburn, Clearing, Woodlawn, South Shore, Oakland, etc.

```{r}
#| fig-height: 6

library(ggridges)

#change label names vv add to density below there
#prettier_labels_PSS <- PSS |>
  mutate(
    Student_Athlete_Code_Better = case_when(
      Student_Athlete_Code == TRUE ~ "Student Athlete", 
      Student_Athlete_Code == FALSE ~ "Non-Student Athlete"
    )
  )

ggplot(
  data = PSS,
  mapping = aes(y = AREA)
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_WHITE),
  color = "red", fill="red", alpha = 0.3
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_AFRICAN_AMERICAN),
  color = "blue", fill="blue", alpha = 0.3
) + labs(
    title = "Trust Level for Police",
    subtitle = "Black vs White population",
    x = "Trust Level (1-100)",
    y = "Area",
    )

ggplot(data = PSS, aes(x=S_RACE_WHITE, color=AREA, fill=AREA)) + geom_density(alpha=.75) + facet_wrap(~AREA)
```

```{r}
ggplot(
  data = PSS,
  mapping = aes(y = AREA)
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_WHITE),
  color = "red", alpha = 0.3
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_HISPANIC),
  color = "blue", alpha = 0.3
) + ggridges::geom_density_ridges(
  mapping = aes(x = S_RACE_AFRICAN_AMERICAN),
  color = "green", alpha = 0.3
) + facet_wrap(~AREA)
```

### Conclusion

I want to check back with my previous workshops that involved EDA and see what the format was there to make this more professional and mirror the academia process for it. It makes it more accessible and preps me for whatever's to come in the future that does this.
